Towards a miniKanren with fair search strategies

### 2019-05-25 Paper/paper.tex
corrections

\newcommand{\disjtwo}{\texttt{disj$_2$}}
\newcommand{\conjtwo}{\texttt{conj$_2$}}
\newcommand{\veryrecursiveo}{\texttt{very-recursive$^o$}}
\newcommand{\appendo}{\texttt{append$^o$}}
\newcommand{\reverso}{\texttt{revers$^o$}}
\newcommand{\repeato}{\texttt{repeat$^o$}}

\newcommand{\DFSi }[0]{DFS$_{i}$}
\newcommand{\DFSf }[0]{DFS$_{f}$}
\newcommand{\DFSbi}[0]{DFS$_{bi}$}
\newcommand{\BFS}[0]{BFS}
\newcommand{\BFSser}[0]{BFS}
\newcommand{\BFSuni}[0]{our BFS}
\newcommand{\BFSimp}[0]{our improved BFS}
\newcommand{\runstar}{run$^*$}

@@

@@

\NOTE[DPF]{Throughout the paper we have search spaces and search strategies.
I can see leaving in the  ``search strategies'', but let's try to
remove virtually all of the uses of ``search'' in front of spaces.}

\NOTE[DPF]{Does this mean that the comment about append goes away.
I really thought that was so clever.  I thought that the way you pointed
out the trivial changes was very clear, so that did not bother me at
all.}

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can be
fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.  The two non-standard
depth-first searches are new. And we present two new, more efficient
and shorter implementations of the standard breadth-first search.
Using quantitative evaluation, we argue that the two new depth-first
searches are competitive alternatives to the standard one, and that
our breadth-first search implementations are more efficient than the
current one.

@@

\NOTE[DPF]{their BFS (or shortened to BFS). We need only say this once if we go with ``our'', which I don't like.
I like ``the new'' or ``the improved'' and ``the current'' }

@@

\DFSi{} is not always the best choice. For instance, it might be less 
desirable for novice miniKanrenuser---understanding implementation details and 
fiddling with clause order is not their first priority. 
There is another reason that miniKanren could use more search strategies than
just \DFSi. In many applications, there does not exist one order that serves all
purposes. For example, a relational dependent type checker contains
clauses for constructors that build data and clauses for eliminators that use
data. When the type checker is generating simple and shallow programs,
the clauses for constructors had better be at the top of the
\conde{} expression.
When performing proof searches for complicated programs, the clauses for 
eliminators had better be at the top of the \conde{} expression. With \DFSi, 
these two uses cannot be efficient at the same time. In fact, to make one use 
efficient, the other one must be more sluggish.

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in two new ways the standard breath-first search (\BFS), a 
search strategy with fair \disj{} and fair \conj{}. We refer to ``the current 
BFS implementation'' (``BFS'' for short) as the implementation by 
\citet{seres1999algebra} and refer to ``our BFS implementation'' (``our BFS'' 
for short) and ``our improved BFS implementation'' (``our improved BFS`` for 
short) as our new ones. We also refer to ``our \BFS{}s'' to both \BFSuni{} and 
\BFSimp{}. We argue that our \BFS{}s are semantically equivalent and formally 
prove that \BFSimp{} and \BFSser{} are semantically equivalent, however, our 
\BFS{}s run faster in all benchmarks and are shorter.
\end{itemize}

\NOTE[DPF]{The citetseres1999algebra must be not broken:
            You can either rewrite the sentence or
            use a tilde called a tie that means don't break the line.
            Sometimes using tilde does not work out elegant enough,
            so the first idea works best.}

@@

In this section, we define fair \disj{}, almost-fair \disj{} and fair \conj{}. 
Before going further into fairness, we give a short review of the terms:
\emph{state}, search \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A \emph{space} is a collection of states. And a \emph{goal} is a 
function from a state to a space.

% Goals check their input states and outpust possibly extended states. A goal 
% might fail for some inputs, in which case the output search space would be 
% empty.

@@

Given the following program, it is natural to expect lists of each
letter to constitute $1/4$ in the query result. \DFSi, TRS2's search
strategy, however, results in many more lists of \texttt{a}s than
lists of other letters. And some letters (e.g. \texttt{c} and
\texttt{d}) are rarely seen. The more clauses the worse the situation.

@@

Under the hood, the \conde{} here is allocating computational resources to 
four trivially different search spaces. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first search space. On the 
contrary, fair \disj{} would allocate resources evenly to each search space.

@@

\begin{defn}[fair \disj{}]
A \disj{} is fair if and only if it allocates computational resources evenly to 
search spaces produced by goals in the same disjunction 
(i.e., clauses in the same \conde).
\end{defn}

\begin{defn}[almost-fair \disj{}]
A \disj{} is almost-fair if and only if it allocates computational resources
so evenly to search spaces produced by goals in the same disjunction that 
the maximal ratio of resources is bounded by a constant.
\end{defn}

@@

Under the hood, the \conde{} and the call to \repeato{} are connected by 
\conj{}. The \conde{} goal outputs a search space including four trivially 
different states. 
Applying the next conjunctive goal, \texttt{(repeato x q)}, produces four 
trivially different search spaces.
In the examples above, all search strategies allocate more computational 
resources to the search space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each search space. For example,

@@

Our solution requires a search strategy with \emph{fair} \conj{} to organize
states in bags in search spaces, where each bag contains finite states, and 
to allocate resources evenly among search spaces derived from states in the 
same bag. It is up to a search strategy designer to decide by what criteria to 
put states in the same bag, and how to allocate resources among search spaces 
related to different bags.

\BFS{} puts states of the same cost in the same bag, and allocates
resources carefully among search spaces related to different bags such
that it produces answers in increasing order of cost. The \emph{cost}
of an answer is its depth in the search tree (i.e., the number of
calls to relations required to find them) \citep{seres1999algebra}. In
the following example, every answer is a list of a list of symbols,
where inner lists in the same outer list are identical. Here the cost of each 
answer is equal
to the length of its inner list plus the length of its outer list. For example,
the cost of \texttt{((a) (a))} is $1 + 2 = 3$.

\NOTE[DPF]{Can you express exactly how all costs are determined?  I see that
for this simple program, your notion of cost makes sense, but how does it
generalize to all costs.}
\NOTE[LKC]{I mean for all answer in the answer list, the answer's inner lists 
are pairwise \texttt{equal?}. It is impossible to see a \texttt{((a) (b))} in 
the answer list.}

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resources evenly to 
search spaces produced from states in the same bag. A bag is a finite 
collection of states. And search strategies with fair \conj{} should represent 
search spaces with possibly unbounded collections of bags. 
\end{defn}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provide a comprehensive description of the 
miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: \texttt{s}s name 
states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name (search) spaces. Fig.~\ref{DFSi-0} shows the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The helper function 
\texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

Balanced interleaving DFS (\DFSbi{}) has an almost-fair \disj{} and unfair 
\conj{}. The implementation of \DFSbi{} differs from 
\DFSi{}'s in the \disj{} macro. We list the new \disj{} with its 
helper in Fig.~\ref{balanced-disj}. When there are one or more disjunctive 
goals, \disj{} builds a balanced binary tree whose leaves are the goals and 
whose nodes are \disjtwo{}s, hence the name of this search strategy. 
The new helper, \texttt{disj+}, takes two additional `arguments'. They 
accumulate goals to be put in the left and right subtrees. The first clause 
handles the case where there is only one goal. In this case, the tree is the 
goal itself. When there are more goals, we partition the list of goals 
into two sublists of roughly equal lengths and recur on the two sublists. We 
move goals to the accumulators in the last clause. As we are moving 
two goals each time, there are two base cases: (1) no goal remains; (2) one 
goal remains. We handle these two new base cases in the second clause and the 
third clause, respectively. In contrast, the \disj{} in \DFSi{} constructs the 
binary tree in a particularly unbalanced form.

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\disjtwo{} (Fig.~\ref{fDFS}). The new \disjtwo{} calls a new and 
fair version of \texttt{append$^\infty$}. \texttt{append$^\infty_{fair}$} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that we haven't swapped
\texttt{s$^\infty$} and \texttt{t$^\infty$}. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \texttt{s$^\infty$} and 
\texttt{t$^\infty$} are ended with a thunk, and that we have swapped them. In 
this case, we construct a new thunk. The new thunk swaps back the two spaces in the
recursive call to \texttt{loop}. This is unnecessary for fairness---we do it to 
produce answers in a more readable order.

@@

\begin{figure}
	\lstinputlisting{Figures/BFSuni.rkt}	
	\caption{implementation of \BFSuni{} (based on \DFSf{})}
	\label{BFSuni}
\end{figure}

\BFS{} has both fair \disj{} and fair \conj{}. Our first implementation 
(\autoref{BFSuni}) is based on \DFSf{} (not \DFSi{}). It is so similar to 
\DFSf{} that we only need to apply two changes: (1) rename 
\texttt{append-map$^\infty$} to \texttt{append-map$^\infty_{fair}$} and (2) 
replace \texttt{append$^\infty$} with \texttt{append$^\infty_{fair}$} in 
\texttt{append-map$^\infty$}'s body.

\NOTE[DPF]{I need clarification: Applying the above changes won't result in 
	``our \BFS,'' but yet a third implementation of \BFS?}
\NOTE[LKC]{Improved.}

\NOTE[DPF]{Why are the two texttt's macros, again?  Now, it hardly matters,
but it is unfriendly to your other co-authors to have us reading long
expressions when a newcommand would suffice.  I read the paper with a pen
marking it up, then I read the paper on the .tex screen.  That's each time
I read it.}

\NOTE[DPF]{You know that ``bag'' is a technical terms.  It means that there
is a structure where each element of the structure has an associated number, 
which tells you have many occurrences of that element are in the list.
For example, the list (a b a c a d c) could be represented like this:
(a$^3$ b$^1$ c$^2$ d$^1$).  If that is not what you mean, you might change the name
to bucket.}

This implementation can be improved in two ways. First, as mentioned in 
\autoref{sec:fairconj}, \BFS{} puts answers in bags and answers of the same 
cost are in the same bag. In the above implementation, however, 
it is not obvious how we manage cost information---the \texttt{car}s of a 
search space have cost 0 (i.e., they are all in the same bag), and every thunk 
indicates an increment in cost. It is even more subtle that 
\texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
respects the cost information. Second, \texttt{append$^\infty_{fair}$} is 
extravagant in memory usage. It makes $O(n+m)$ new \texttt{cons} cells every 
time, where $n$ and $m$ are the ``length''s of input spaces. \DFSf{} also has 
the space extravagance.

@@

\NOTE[DPF]{Let's use one more line and far less horizontal space
by use 3 linefeeds.  They will still align nicely.}
\NOTE[LKC]{fixed}

@@

\begin{figure}
	\lstinputlisting{Figures/BFSimp.rkt}	
	\caption{New and changed functions in our BFS that implements pure 
	features}
	\label{BFSimp}
\end{figure}

% As mentioned in \autoref{sec:fairconj}, \BFS{} puts answers in bags and 
% answers of the same cost are in the same bag. 
% In the easy implementation, cost information is recorded subtly---the 
% \texttt{car}s of a search space have cost 0 (i.e. they are all in the same 
% bag),
% and every thunk indicates an increment in cost. It is even more subtle that
% \texttt{append$^\infty_{fair}$} and the \texttt{append-map$^\infty_{fair}$} 
% respects the cost information. We make these facts more obvious by changing 
% the type of search space, modifying related function definitions, and 
% introducing a few more functions.

\NOTE[DPF]{Font Space like in the code comments.}

We make the cost information more clear by changing the Space type, 
modifying related function definitions, and introducing a few more functions.

The new type is a pair whose \texttt{car} is a list of answers (the bag), and 
whose \texttt{cdr} is either \texttt{\#{}f} or a thunk returning a search 
space. A falsy \texttt{cdr} means the search space is obviously finite.

We list functions related to the pure subset in Fig.~\ref{BFSimp}. The first 
three functions are search space constructors. \texttt{none} makes an empty 
search space; \texttt{unit} makes a space from one answer; and \texttt{step} 
makes a space from a thunk. The remaining functions are as before. We compare 
these functions with \BFSser{} in our proof.

Luckily, the change in \texttt{append$^\infty_{fair}$} also fixes the miserable 
space extravagance---the use of \texttt{append} helps us to reuse the first bag 
of \texttt{t$^\infty$}.

\citet{kiselyov2005backtracking} have demonstrated that a \emph{MonadPlus} 
hides in implementations of logic programming system. \BFSimp{} is not an 
exception: \texttt{append-map$^\infty_{fair}$} is like \texttt{bind}, but takes 
arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\texttt{append$^\infty_{fair}$} correspond to \texttt{mzero}, \texttt{unit}, and 
\texttt{mplus}, respectively.

\begin{figure}
	\lstinputlisting{Figures/BFSimp-cont.rkt}	
	\caption{New and changed functions in our BFS that implement impure 
		features}
	\label{BFSimp-cont}
\end{figure}

\NOTE[DPF]{ks --> sk and kf --> fk and you don't need to say f and s
mean success continuation and failure continuation.}

Functions implementing impure features are in Fig.~\ref{BFSimp-cont}. The 
first function, \texttt{elim}, takes a space \texttt{s$^\infty$} and two 
continuations \texttt{kf} and \texttt{ks}. When \texttt{s$^\infty$} contains 
no answers, it calls \texttt{kf} with no argument. Otherwise, it calls 
\texttt{ks} with the first answer and the rest of the space. Here  `f' means 
`fail' and `s' means `succeed'. This function is similar to an eliminator of 
search spaces, hence the name. The remaining functions are as before.

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both our BFS and BFS. (See supplements for the formal proof.) To 
compare efficiency, we 
translate \BFS{}'s Haskell code into Racket (See supplements for the 
translated code). The translation is direct due to the similarity of the two 
relational programming systems. The translated code is longer than \BFSimp{}. 
And it runs slower in all benchmarks. Details about differences in efficiency 
are in \autoref{sec:quan}.
% \autoref{compare-efficiency}

@@

We analyze the definitions of fair \disj{} and fair \conj{}, then propose a 
new definition of fair \conj{}. Our definition is orthogonal with completeness.

We devise two new search strategies (i.e. balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise two new 
implementations of \BFS. These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}.

Our quantitative evaluation shows that \DFSbi{} is a competitive 
alternative to \DFSi{}, the current miniKanren search strategy,
and that \BFS{} is less practical.
% and that \BFSser{} is less practical than other breadth-first strategies.

Our improved \BFS{} is semantically equivalent to the original 
one. But, ours is shorter and performs better in comparison with a 
straightforward translation of their Haskell code.

>>>>

\newcommand{\disjtwo}{\texttt{disj$_2$}}
\newcommand{\conjtwo}{\texttt{conj$_2$}}
\newcommand{\veryrecursiveo}{\texttt{very-recursive$^o$}}
\newcommand{\appendo}{\texttt{append$^o$}}
\newcommand{\reverso}{\texttt{revers$^o$}}
\newcommand{\repeato}{\texttt{repeat$^o$}}
\newcommand{\runstar}{\texttt{run$^*$}}
\newcommand{\Space}{\texttt{Space}}
\newcommand{\appendInf}{\texttt{append$^\infty$}}
\newcommand{\appendInfFair}{\texttt{append$^\infty_{fair}$}}
\newcommand{\appendMapInf}{\texttt{append-map$^\infty$}}
\newcommand{\appendMapInfFair}{\texttt{append-map$^\infty_{fair}$}}
\newcommand{\sInf}{\texttt{s$^\infty$}}
\newcommand{\tInf}{\texttt{t$^\infty$}}

\newcommand{\DFSi }[0]{DFS$_{i}$}
\newcommand{\DFSf }[0]{DFS$_{f}$}
\newcommand{\DFSbi}[0]{DFS$_{bi}$}
\newcommand{\BFS}[0]{BFS}
\newcommand{\BFSser}[0]{BFS}
\newcommand{\BFSimp}[0]{the improved BFS}

@@

\NOTE[DPF]{Throughout the paper we have search spaces and search strategies.
I can see leaving in the  ``search strategies'', but let's try to
remove virtually all of the uses of ``search'' in front of spaces.}
\NOTE[LKC]{Corrected.}

\NOTE[DPF]{Does this mean that the comment about append goes away.
I really thought that was so clever.  I thought that the way you pointed
out the trivial changes was very clear, so that did not bother me at
all.}
\NOTE[LKC]{Yes, the comment about append will go away if we remove our improved 
BFS. Then, let's don't remove it, i.e. have two new BFS implementations at the 
same time.}
\NOTE[LKC]{I think having two new BFSs is a bit confusing. So I brand the 
new and easy/unimproved one as a ``stepping-stone'' and only mention it in BFS 
section.}

@@

We describe fairness levels in disjunction and conjunction
implementations.  Specifically, a disjunction implementation can be
fair, almost-fair, or unfair. And a conjunction implementation can be
fair or unfair.  We compare the fairness level of four search
strategies: the standard miniKanren interleaving depth-first search,
the balanced interleaving depth-first search, the fair depth-first
search, and the standard breadth-first search.  The two non-standard
depth-first searches are new. And we present a new, more efficient
and shorter implementation of the standard breadth-first search.
Using quantitative evaluation, we argue that a new depth-first searches is a
competitive alternative to the standard one, and that our improved 
breadth-first search implementation is more efficient than the
current one.

@@

\NOTE[DPF]{their BFS (or shortened to BFS). We need only say this once if we go 
with ``our'', which I don't like.
I like ``the new'' or ``the improved'' and ``the current'' }
\NOTE[LKC]{Corrected. I pick ``the improved''.}

@@

\DFSi{} is not always the best choice. For instance, it might be less 
desirable for novice miniKanren users---understanding implementation details 
and fiddling with clause order is not their first priority. 
There is another reason that miniKanren could use more search strategies than
just \DFSi. In many applications, there does not exist one order that serves all
purposes. For example, a relational dependent type checker contains
clauses for constructors that build data and clauses for eliminators that use
data. When the type checker is generating simple and shallow programs,
the clauses for constructors had better be at the top of the
\conde{} expression.
When performing proof searches for complicated programs, the clauses for 
eliminators had better be at the top of the \conde{} expression. With \DFSi, 
these two uses cannot be efficient at the same time. In fact, to make one use 
efficient, the other one must be more sluggish.

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in a new way the standard breath-first search (\BFS) by 
\citet{seres1999algebra}, a search strategy with fair \disj{} and fair \conj{}. 
We refer to ``\BFSser{}'' as their implementation and ``the improved 
BFS'' as our new one. We formally prove that the two implementations are 
semantically equivalent, however, \BFSimp{} runs faster in all benchmarks and 
is shorter.
%	\item implement in a new way the standard breath-first search (\BFS), a 
%search strategy with fair \disj{} and fair \conj{}. We refer to ``the current 
%BFS implementation'' (``BFS'' for short) as the one by 
%\citet{seres1999algebra} 
%and refer to ``our improved BFS implementation'' (``the improved BFS'' for 
%short) as the new one. We formally prove that \BFSimp{} and \BFSser{} are 
%semantically equivalent, however, \BFSimp{} runs faster in all benchmarks and 
%is shorter.
\end{itemize}

\NOTE[DPF]{The citetseres1999algebra must be not broken:
            You can either rewrite the sentence or
            use a tilde called a tie that means don't break the line.
            Sometimes using tilde does not work out elegant enough,
            so the first idea works best.}
\NOTE[LKC]{I have rewritten the whole item.}

@@

In this section, we define fair \disj{}, almost-fair \disj{} and fair \conj{}. 
Before going further into fairness, we give a short review of the terms:
\emph{state}, \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A \emph{space} is a collection of states. And a \emph{goal} is a 
function from a state to a space.

@@

Given the following program, it is natural to expect lists of each
letter to constitute $1/4$ in the answer list. \DFSi, TRS2's search
strategy, however, results in many more lists of \texttt{a}s than
lists of other letters. And some letters (e.g. \texttt{c} and
\texttt{d}) are rarely seen. The more clauses the worse the situation.

@@

Under the hood, the \conde{} here is allocating computational resources to 
four trivially different spaces. The unfair \disj{} in 
\DFSi{} allocates many more resources to the first space. On the 
contrary, fair \disj{} would allocate resources evenly to each space.

@@

\begin{defn}[fair \disj{}]
A \disj{} is fair if and only if it allocates computational resources evenly to 
spaces produced by goals in the same disjunction 
(i.e., clauses in the same \conde).
\end{defn}

\begin{defn}[almost-fair \disj{}]
A \disj{} is almost-fair if and only if it allocates computational resources
so evenly to spaces produced by goals in the same disjunction that 
the maximal ratio of resources is bounded by a constant.
\end{defn}

@@

Under the hood, the \conde{} and the call to \repeato{} are connected by 
\conj{}. The \conde{} goal outputs a space including four trivially 
different states. 
Applying the next conjunctive goal, \texttt{(repeato x q)}, produces four 
trivially different spaces.
In the examples above, all search strategies allocate more computational 
resources to the space of \texttt{a}. On the contrary, fair \conj{} 
would allocate resources evenly to each space. For example,

@@

Our solution requires a search strategy with \emph{fair} \conj{} to organize
states in buckets in spaces, where each bucket contains finite states, 
and 
to allocate resources evenly among spaces derived from states in the 
same bucket. It is up to a search strategy designer to decide by what criteria 
to 
put states in the same bucket, and how to allocate resources among spaces 
related to different buckets.

\BFS{} puts states of the same cost in the same bucket, and allocates
resources carefully among spaces related to different buckets such
that it produces answers in increasing order of cost. The \emph{cost}
of an answer is its depth in the search tree (i.e., the number of
calls to relational definitions required to find the answer) 
\citep{seres1999algebra}. In the above examples, the cost of each answer is 
equal to their lengths because we need to apply \repeato{} $n$ times to find an 
answer of length $n$. In the following example, every answer is a list of a 
list of symbols, where inner lists in the same outer list are identical. Here 
the cost of each answer is equal to the length of its inner lists plus the 
length of its outer list. For example, the cost of \texttt{((a) (a))} is $1 + 2 
= 3$.

\NOTE[DPF]{Can you express exactly how all costs are determined?  I see that
for this simple program, your notion of cost makes sense, but how does it
generalize to all costs.}
\NOTE[LKC]{Corrected.}

@@

\begin{defn}[fair \conj{}]
A \conj{} is fair if and only if it allocates computational resources evenly to 
spaces produced from states in the same bucket. A bucket is a finite 
collection of states. And search strategies with fair \conj{} should represent 
spaces with possibly unbounded collections of buckets. 
\end{defn}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provide a comprehensive description of the miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search strategies. We follow some conventions to name variables: 
\texttt{s}s name states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name (search) spaces. Fig.~\ref{DFSi-0} shows 
the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting spaces with \appendInf{}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting space. The helper function 
\appendMapInf{} applies its input goal to states 
in its input space and composes the resulting spaces. It reuses 
\appendInf{} for space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

Balanced interleaving DFS (\DFSbi{}) has an almost-fair \disj{} and unfair 
\conj{}. The implementation of \DFSbi{} differs from 
\DFSi{}'s in the \disj{} macro. When there are one or more disjunctive 
goals, the new \disj{} builds a balanced binary tree whose leaves are the goals 
and 
whose nodes are \disjtwo{}s, hence the name of this search strategy. 
In contrast, the \disj{} in \DFSi{} constructs the binary tree in a 
particularly unbalanced form.
We list the new \disj{} with its helper in Fig.~\ref{balanced-disj}.
The new helper, \texttt{disj+}, takes two additional `arguments'. They 
accumulate goals to be put in the left and right subtrees. The first clause 
handles the case where there is only one goal. In this case, the tree is the 
goal itself. When there are more goals, we partition the list of goals 
into two sublists of roughly equal lengths and recur on the two sublists. We 
move goals to the accumulators in the last clause. As we are moving 
two goals each time, there are two base cases: (1) no goal remains; (2) one 
goal remains. We handle these two new base cases in the second clause and the 
third clause, respectively.

@@

Fair DFS (\DFSf) has fair \disj{} and unfair \conj{}. The 
implementation of \DFSf{} differs from \DFSi{}'s in 
\disjtwo{} (Fig.~\ref{fDFS}). The new \disjtwo{} calls a new and 
fair version of \appendInf{}. \appendInfFair{} 
immediately 
calls 
its helper, \texttt{loop}, with the first argument, \texttt{s?}, set to 
\texttt{\#{}t}, which indicates that we haven't swapped
\sInf{} and \tInf{}. The swapping 
happens at 
the third \texttt{cond} clause in the helper, where \texttt{s?} is updated 
accordingly. The first two \texttt{cond} clauses essentially copy the 
\texttt{car}s and stop recursion when one of the input spaces is obviously 
finite. The third clause, as we mentioned above, is just for swapping. When the 
fourth and last clause runs, we know that both \sInf{} and 
\tInf{} are ended with a thunk, and that we have swapped them. In 
this case, we construct a new thunk. The new thunk swaps back the two spaces in 
the
recursive call to \texttt{loop}. This is unnecessary for fairness---we do it to 
produce answers in a more readable order.

@@

\begin{figure}
	\lstinputlisting{Figures/BFSuni.rkt}	
	\caption{stepping-stone toward \BFSimp{} (based on \DFSf{})}
	\label{BFSuni}
\end{figure}

\BFS{} has both fair \disj{} and fair \conj{}. Our first BFS implementation 
serves as a ``stepping-stone'' toward \BFSimp{}. It is so similar to \DFSf{} 
(not \DFSi{}) that we only need to apply two changes: (1) rename 
\appendMapInf{} 
to \appendMapInfFair{} and (2) replace \appendInf{} with \appendInfFair{} in 
\appendMapInfFair{}'s body.

\NOTE[DPF]{I need clarification: Applying the above changes won't result in 
	``our \BFS,'' but yet a third implementation of \BFS?}
\NOTE[LKC]{If you still need clarification, please add a new note below. 
Otherwise, you might delete these two notes.}

\NOTE[DPF]{Why are the two texttt's macros, again?  Now, it hardly matters,
but it is unfriendly to your other co-authors to have us reading long
expressions when a newcommand would suffice.  I read the paper with a pen
marking it up, then I read the paper on the .tex screen.  That's each time
I read it.}
\NOTE[LKC]{I am sorry about that. Thank you for telling me! I have added many 
more macros and have removed many uses of texttt.}

\NOTE[DPF]{You know that ``bag'' is a technical terms.  It means that there
is a structure where each element of the structure has an associated number, 
which tells you have many occurrences of that element are in the list.
For example, the list (a b a c a d c) could be represented like this:
(a$^3$ b$^1$ c$^2$ d$^1$).  If that is not what you mean, you might change the 
name to bucket.}
\NOTE[LKC]{That is not what I mean. I have changed the name to bucket.}

This implementation can be improved in two ways. First, as mentioned in 
\autoref{sec:fairconj}, \BFS{} puts answers in buckets and answers of the same 
cost are in the same bucket. In the above implementation, however, 
it is not obvious how we manage cost information---the \texttt{car}s of a 
space have cost 0 (i.e., they are all in the same bucket), and every 
thunk indicates an increment in cost. It is even more subtle that 
\appendInfFair{} and the \appendMapInfFair{} respects the cost information. 
Second, \appendInfFair{} is extravagant in memory usage. It makes $O(n+m)$ new 
\texttt{cons} cells every time, where $n$ and $m$ are the sizes of 
the first buckets of two input spaces. \DFSf{} also has the space extravagance.

\NOTE[LKC]{I rewrite the second last sentence in the above paragraph such that 
it better corresponds to the comment about `append'.}

@@

@@

\begin{figure}
	\lstinputlisting{Figures/BFSimp.rkt}	
	\caption{New and changed functions in \BFSimp{} that implements pure 
	features}
	\label{BFSimp}
\end{figure}

\NOTE[DPF]{Font Space like in the code comments.}
\NOTE[LKC]{corrected}

We make the cost information more clear by changing the \texttt{Space} type, 
modifying related function definitions, and introducing a few more functions.

The new type of \texttt{Space} is a pair whose \texttt{car} is a list of 
answers (the bucket), and whose \texttt{cdr} is either \texttt{\#{}f} or a 
thunk returning a space. A falsy \texttt{cdr} means the space is obviously 
finite.

We list functions related to the pure subset in Fig.~\ref{BFSimp}. The first 
three functions are space constructors. \texttt{none} makes an empty space; 
\texttt{unit} makes a space from one answer; and \texttt{step} makes a space 
from a thunk. The remaining functions are as before. We compare these functions 
with \BFSser{} in our proof.

Luckily, the change in \appendInfFair{} also fixes the miserable space 
extravagance---the use of \texttt{append} helps us to reuse the first 
bucket of \tInf{}.

\citet{kiselyov2005backtracking} have demonstrated that a \emph{MonadPlus} 
hides in implementations of logic programming systems. \BFSimp{} is not an 
exception: \appendMapInfFair{} is like \texttt{bind}, but takes 
arguments in reversed order; \texttt{none}, \texttt{unit}, and 
\appendInfFair{} correspond to \texttt{mzero}, \texttt{unit}, and 
\texttt{mplus}, respectively.

\begin{figure}
	\lstinputlisting{Figures/BFSimp-cont.rkt}	
	\caption{New and changed functions in \BFSimp{} that implement impure 
		features}
	\label{BFSimp-cont}
\end{figure}

\NOTE[DPF]{ks --> sk and kf --> fk and you don't need to say f and s
mean success continuation and failure continuation.}
\NOTE[LKC]{corrected}

Functions implementing impure features are in Fig.~\ref{BFSimp-cont}. The 
first function, \texttt{elim}, takes a space \sInf{} and two 
continuations \texttt{fk} and \texttt{sk}. When \sInf{} contains 
no answers, it calls \texttt{fk} with no argument. Otherwise, it calls 
\texttt{sk} with the first answer and the rest of the space. This function is 
similar to an eliminator of spaces, hence the name. The remaining 
functions are as before.

@@

In this subsection, we compare the pure subset of \BFSimp{} with \BFSser{}. We 
focus on the pure subset because \BFSser{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both \BFSimp{} and BFS. (See supplements for the formal proof.) To 
compare efficiency, we translate \BFSser{}'s Haskell code into Racket (See 
supplements for the translated code). The translation is direct due to the 
similarity of the two relational programming systems. The translated code is 
longer than \BFSimp{}. And it runs slower in all benchmarks. Details about 
differences in efficiency are in \autoref{sec:quan}.

@@

We analyze the definitions of fair \disj{} and fair \conj{}, then propose a 
new definition of fair \conj{}. Our definition of fair \conj{} is orthogonal 
with completeness.

We devise two new search strategies (i.e. balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS. These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}.

Our quantitative evaluation shows that \DFSbi{} is a competitive 
alternative to \DFSi{}, the current miniKanren search strategy,
and that \BFSimp{} is more practical than \BFSser{}.

% \BFSimp{}
We proof formally that \BFSimp{} is semantically equivalent to \BFSser{}. But, 
\BFSimp{} is shorter and performs better in comparison with a straightforward 
translation of their Haskell code.

We acknowledge that we are making large claims about efficiency given that we 
have just a few benchmarks, but our intuition is based on sound principles. ...
